---
title: "P8105 Homework 5"
author: "Diana Hernandez"
date: "2023-11-15"
output: github_document
---

Load libraries and settings.

```{r}
library(tidyverse)
library(rvest)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal())

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

Import the data.

```{r}
homicide_data =
  read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")
```

The raw data has `r ncol(homicide_data)` variables and `r nrow(homicide_data)` observations. Some of the key variables include victim name, victim age, victim race, victim sex, city and state of homicide, and disposition. 

Create `city_state` variable and create summary for the total number of homicides and the number of unsolved homicides.

```{r}
homicide_tidy =
  homicide_data |>
   mutate(
      city_state = str_c(city, state, sep = ", ")
    ) |>
  filter(!city_state == "Tulsa, AL")

homicide_df =
  homicide_tidy |>
  group_by(city_state) |>
  summarize(n_obs = n())

unsolved_homicide_df =
  homicide_tidy |>
  filter(disposition %in% c("Closed without arrest", "Open/No arrest")) |>
  group_by(city_state) |>
  summarize(n_obs = n())
```

Use the `prop.test` function for `city_state` == "Baltimore,MD".

```{r}
homicide_prop_test = function(x) {

n_homicide =  
  homicide_df |>
  filter(city_state == x) |>
  pull(n_obs)
  
n_unsolved =
  unsolved_homicide_df |>
  filter(city_state == x) |>
  pull(n_obs)
  
  
result_tidy = broom::tidy(prop.test(x = n_unsolved, n = n_homicide))

}

homicide_prop_test("Baltimore, MD")
```

Use `prop.test` for all cities. I could not figure out how to keep `city_state` variable, but I see what I could have done different by looking at Jeff's solution for Problem 1.

```{r}
cities = unique(homicide_df$city_state)

prop_test_output = map(cities, homicide_prop_test)

prop_test_df =
  bind_rows(prop_test_output)

prop_test_df |>
  select(estimate, conf.low, conf.high)
```

A plot that shows the estimates and CIs for each city. Cities are depicted as number since I could not figure out how to keep the `city_state` variable.

```{r}
prop_test_df |>
  mutate(city_number = row_number(),
         city_number = as.factor(city_number),
         city_number = fct_reorder(city_number, estimate)) |>
  ggplot(aes(x = city_number, y = estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Estimates and CIs for each city",
    x = "City ID",
    y = "Estimate"
  )
```

# Problem 2

Import the data.

```{r}
data_path = "./data/" 

import_csv = function(x) {
  read_csv(x) |>
  mutate(filename = x)
}

con_exp = list.files(data_path, pattern = "\\.csv$", full.names = T)

con_exp_df = map(con_exp, import_csv)

final_con_exp_data = bind_rows(con_exp_df)
```

Data wrangling.

```{r}
tidy_con_exp_data =
  final_con_exp_data |>
  mutate(
    subject_id = row_number(),
    control_arm = ifelse(subject_id < 11, "Control", "Experimental")
    ) |>
  select(subject_id, control_arm, everything(), -filename) |>
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "observations"
  )
```

Spaghetti plot.

```{r}
tidy_con_exp_data |> 
  ggplot(aes(x = week, y = observations, group = subject_id, color = subject_id)) +
  geom_line() +
  labs(
    title = "Observations over time by subject",
    x = "Week",
    y = "Observations",
    color = "Subject ID"
    )

tidy_con_exp_data |> 
  ggplot(aes(x = week, y = observations, group = subject_id, color = control_arm)) +
  geom_line() +
  labs(
    title = "Observations over time by subject and control arm",
    x = "Week",
    y = "Observations",
    color = "Control Arm"
    )
```

I created two graphs: one to compare each individual subject and another to compare control and experimental groups. Since subjects in the experimental group are 11-20, you can see that their observations over time are higher than those subjects in the control group (1-10). 

# Problem 3

Create function. Set Mu = 0. 

```{r}
one_samp_ttest = function(n = 30, sigma = 5, mu = 0) {
  
sample_data = 
  tibble(
      rnorm(n, mean = mu, sd = sigma),
    ) 
  
broom::tidy(t.test(sample_data, mu = 0, alternative = "two.sided"))
  
}
```

Generate 5000 datasets from the model.

```{r}
ttest_results_df = 
  expand_grid(
    sample_size = 30,
    iter = 1:5000
  ) |> 
  mutate(
    estimate_df = map(sample_size, one_samp_ttest)
  ) |> 
  unnest(estimate_df)

ttest_results_df |>
  select(estimate, p.value)
```

Perform `t.test` for μ = {1,2,3,4,5,6}.

```{r}
t_test_multiple_means = function(n = 30, sigma = 5, population_mean) {

  perform_t_test = function(mu) {
    sample_data = 
      tibble(
        rnorm(n, mean = population_mean, sd = sigma),
    ) 
    
  t_test_result = broom::tidy(t.test(sample_data, mu = mu, alternative = "two.sided"))

  }
  
  t_test_results = map(population_mean, perform_t_test) 
  
}

population_mean = c(1, 2, 3, 4, 5)

multiple_mu_results = t_test_multiple_means(n = 30, sigma = 5, population_mean)
```

Generate 5000 datasets for μ = {1,2,3,4,5,6}.

```{r}
results_df = 
  expand_grid(
    sample_size = 30,
    population_mean = c(1, 2, 3, 4, 5),
    iter = 1:5000
  ) |> 
  mutate(
    estimate_df = 
      map2(sample_size, population_mean, \(n, population_mean) t_test_multiple_means(n = 30, sigma = 5, population_mean))
  ) |>
  unnest(estimate_df) |>
  unnest(estimate_df)
```

Create plots.

Power and effect size plot.

```{r}
power_plot = 
  results_df |>
  select(population_mean, estimate, p.value) |>
  mutate(
    reject_null = if_else(p.value < 0.05, 1, 0)
  ) |>
  group_by(population_mean, reject_null) |>
  summarize(
    n_obs = n()) |>
  mutate(props = sum(reject_null) / n_obs) |>
  filter(reject_null == 1) |>
  ggplot(aes(x = population_mean, y = props)) +
  geom_point() +
  labs(
    title = "Effect size and power",
    x = "True value of μ",
    y = "Power of test"
  )

power_plot
```

There is no clear association between effect size and power. 


Average estimate of μ and true value of μ

```{r}
estimate_plot =
  results_df |>
  group_by(population_mean) |>
  summarize(
    mean_mu = mean(estimate)
  ) |>
  ggplot(aes(x = population_mean, y = mean_mu)) +
  geom_point() +
  labs(
    title = "Average estimate of μ and true value of μ.",
    x = "True value of μ",
    y = "Average estimate of μ"
  )

estimate_plot
```

The sample average of μ̂ across tests is equal to the true value of μ.


Only in samples for which the null was rejected.

```{r}
rejected_null_plot =
  results_df |>
  filter(p.value < 0.05) |>
  group_by(population_mean) |>
  summarize(
    mean_mu = mean(estimate)
  ) |>
  ggplot(aes(x = population_mean, y = mean_mu)) +
  geom_point() +
  labs(
    title = "Average estimate of μ and true value of μ.",
    x = "True value of μ",
    y = "Average estimate of μ"
  )

rejected_null_plot
```

The sample average of μ̂ across tests for which the null is rejected approximately equal to the true value of μ likely due to the high power of the test.
